---
title: "R Notebook"
output: html_notebook
---

This importation package runs a model to estimate importations of SARS-CoV-2 into airports globally. The current package includes data to estimate importations into U.S. airports.

```{r setup}

library(covidImportation)
packageVersion("covidImportation")


locs_of_interest <- c("GUM","VIR","ASM","MNP")
states_of_interest <- c("GU", "VI", "AS", "MP")
regioncode="usa_territories"

```

    

# Function to get data for Territories
```{r}

get_oag_travel <- function(destination=c("CA"),
                           destination_type="state",
                           dest_country="USA",
                           dest_aggr_level="city"){
    
    # # check if destination is in the USA
    # # -- only US aggregated data are available in the package
    # if (dest_country!="USA"){
    #     print("Only aggregated, averaged travel data for travel into the USA are included in this package. 
    #           For other countries, provide your own data and use 'get_oag_travel_fulldata' or contact Shaun Truelove (shauntruelove@jhu.edu).")
    #     return(NA)
    # }
    
    # Load the data
    # load_travel_dat <- function(dest_country){
    #     env <- new.env()
    #     dest_data <- data(list=tolower(paste0(dest_country, "_oag_aggr_travel")), 
    #                       package = "covidImportation", 
    #                       envir = env)[1]
    #     return(env[[dest_data]])
    # }
    # dest_data <- load_travel_dat(dest_country)
    
    dest_data <- readr::read_csv("data_other/us_terr_oag_aggr_travel.csv")
    
    # subset to the destination of interest
    dest_data <- dest_data %>% as.data.frame() %>% 
        dplyr::filter(get(paste0("arr_",destination_type)) %in% destination) #%>%
        #select(-rr_state)
    
    return(dest_data)
}


```

    

    
## Setup the data for the importation model.
    
```{r, warning=FALSE, message=FALSE}

# 
# setup_res <- covidImportation::setup_importations(dest=locs_of_interest,
#                                dest_type=c("country"), #,"city","airport", "country"),
#                                dest_country=locs_of_interest,
#                                dest_aggr_level=c("airport"), #, "city", "state", "country", "metro"),
#                                first_date = ISOdate(2019,12,1),
#                                last_date = Sys.time(),
#                                update_case_data=TRUE,
#                                case_data_dir = "data/case_data",
#                                output_dir = file.path("importation",regioncode),
#                                check_saved_data=TRUE,
#                                save_case_data=TRUE,
#                                get_travel=TRUE,
#                                n_top_dests=Inf, 
#                                travel_dispersion=3,
#                                param_list=list(incub_mean_log=log(5.89),
#                                                incub_sd_log=log(1.74),
#                                                inf_period_nohosp_mean=15,
#                                                inf_period_nohosp_sd=5,
#                                                inf_period_hosp_shape=0.75,
#                                                inf_period_hosp_scale=5.367,
#                                                p_report_source=c(0.05, 0.25),
#                                                shift_incid_days=-10,
#                                                delta=1))



dest=locs_of_interest
                               dest_type=c("country")#"city","airport", "country"),
                               dest_country=locs_of_interest
                               dest_aggr_level=c("airport")# "city", "state", "country", "metro"),
                               first_date = ISOdate(2019,12,1)
                               last_date = Sys.time()
                               update_case_data=TRUE
                               case_data_dir = "data/case_data"
                               output_dir = file.path("importation",regioncode)
                               check_saved_data=TRUE
                               save_case_data=TRUE
                               get_travel=TRUE
                               n_top_dests=Inf 
                               travel_dispersion=3
                               param_list=list(incub_mean_log=log(5.89),
                                               incub_sd_log=log(1.74),
                                               inf_period_nohosp_mean=15,
                                               inf_period_nohosp_sd=5,
                                               inf_period_hosp_shape=0.75,
                                               inf_period_hosp_scale=5.367,
                                               p_report_source=c(0.05, 0.25),
                                               shift_incid_days=-10,
                                               delta=1)


# setup_importations <- function(dest="UT",
#                                dest_type=c("state"), #,"city","airport", "country"),
#                                dest_country="USA",
#                                dest_aggr_level=c("airport"), #, "city", "state", "country", "metro"),
#                                first_date = ISOdate(2019,12,1),
#                                last_date = Sys.time(),
#                                update_case_data=TRUE,
#                                case_data_dir = "data/case_data",
#                                output_dir = file.path("output", paste0(paste(dest, collapse="+"),"_", as.Date(Sys.Date()))),
#                                check_saved_data=TRUE,
#                                save_case_data=TRUE,
#                                get_travel=TRUE,
#                                n_top_dests=Inf, 
#                                travel_dispersion=3,
#                                param_list=list(incub_mean_log=log(5.89),
#                                                incub_sd_log=log(1.74),
#                                                inf_period_nohosp_mean=15,
#                                                inf_period_nohosp_sd=5,
#                                                inf_period_hosp_shape=0.75,
#                                                inf_period_hosp_scale=5.367,
#                                                p_report_source=c(0.05, 0.25),
#                                                shift_incid_days=-10,
#                                                delta=1),
#                                pop_additional = NULL){

  ## Create needed directories
  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

  ## DATA
  ## ~ Incidence data
  incid_data_list <- get_incidence_data(first_date = first_date,
                                        last_date = last_date,
                                        update_case_data = update_case_data,
                                        case_data_dir = case_data_dir,
                                        check_saved_data = check_saved_data,
                                        save_data = save_case_data)
  
  incid_data <- incid_data_list$incid_data %>% dplyr::filter(source != "USA") # get rid of generic "USA"
  incid_data <- incid_data %>% rename(incid_est = cases_incid)
  jhucsse <- incid_data_list$jhucsse_case_data
  jhucsse_state <- incid_data_list$jhucsse_case_data_state
  
  
  ## ~ Travel Data
  ## if travel data exists load it, otherwise download it
  if(get_travel) {
    travel_data_monthly <- get_oag_travel(destination=dest,
                                          destination_type=dest_type,
                                          dest_country=dest_country,
                                          dest_aggr_level=dest_aggr_level) %>% as.data.frame()
    travel_data_monthly <- travel_data_monthly %>%
      dplyr::mutate(t_year=2020) %>%
      dplyr::rename(source = dep_loc_aggr)
    travel_data_monthly$destination <- travel_data_monthly[,paste0("arr_", dest_aggr_level),drop=T]
  } else{
    travel_data_monthly <- paste0("data/", paste(dest, collapse = "+"), "-",
                                  dest_aggr_level, "_oag_20172019.csv") %>%
      readr::read_csv(na=c(""," ","NA"))%>%
      dplyr::mutate(t_year=2020) %>%
      dplyr::mutate(travelers=ifelse(t_month == "01" & dep_country=="CHN",
                              # Increase travel for Chinese New Year
                              travelers*1.6, travelers)) %>%
      dplyr::rename(source = dep_loc_aggr)
    travel_data_monthly$destination <- travel_data_monthly[,paste0("arr_", dest_aggr_level),drop=T]
  }
  
  ## monthly average totals into destinations
  travel_mean <- travel_data_monthly %>%
    dplyr::group_by(destination, t_month) %>%
    dplyr::summarise(travelers = sum(travelers_mean,na.rm=TRUE)) %>%
    dplyr::group_by(destination) %>%
    dplyr::summarise(travelers = mean(travelers)) %>%
    dplyr::arrange(desc(travelers))
  
  
  
  # Destinations to keep
  dests_keep <- travel_mean$destination[seq_len(min(c(nrow(travel_mean), n_top_dests)))]
  travel_data_monthly <- travel_data_monthly %>% dplyr::filter(destination %in% dests_keep)
  
  ## Travel data
  ##  - Get daily for merging purposes
  travel_data_daily <- covidImportation:::make_daily_travel(travel_data_monthly, travel_dispersion=3)
  
  ## ~ Population Data
  data(pop_data, package="covidImportation")
  pop_data <- bind_rows(pop_data, pop_additional)
  
  
  ## ~~ First Check that the variables match up
  # Check that incidence data does not have duplicates
  incid_dups <- sum(incid_data %>%
                      dplyr::mutate(source_t = paste(source, t)) %>%
                      dplyr::mutate(dup_entry=duplicated(source_t)) %>%
                      dplyr::pull(dup_entry))
  
  if(sum(incid_dups)>0){
    dup_entry <- incid_data %>%
      dplyr::mutate(source_t = paste(source, t)) %>%
      dplyr::mutate(dup_entry=duplicated(source_t)) %>%
      dplyr::filter(dup_entry) %>% dplyr::pull(source_t)
    warning("There are duplicate entries in the incidence data.")
  }
  ## Check travel data
  travel_dups <- travel_data_daily %>%
    dplyr::mutate(source_dest_t = paste(source, destination, t),
           dup_entry=duplicated(source_dest_t)) %>%
    dplyr::pull(dup_entry) %>%
    sum()
  if(sum(travel_dups)>0){
    warning("There are duplicate entries in the travel data.")
  }
  ## Check Population data
  pop_dups <- pop_data %>%
    dplyr::mutate(dup_entry=duplicated(source)) %>%
    dplyr::pull(dup_entry) %>%
    sum()
  if(sum(pop_dups)>0){
    warning("There are duplicate entries in the population data.")
  }
  # we really just need to make sure there are travel data and pop data for all source locations with incidence
  # incid_sources <- sort(unique(incid_data$source))
  # travel_sources <- sort(unique(travel_data_daily$source))
  # pop_sources <- sort(unique(pop_data$source))
  # incid_sources[!(incid_sources %in% travel_sources)]
  # incid_sources[!(incid_sources %in% pop_sources)]
  
  ## ~~ Merge it all
  input_data <- covidImportation:::make_input_data(incid_data, travel_data_daily, pop_data,
                                shift_incid_days=param_list$shift_incid_days,
                                dest_aggr_level=dest_aggr_level) %>%
    dplyr::mutate(p_report_source=ifelse(source=="Hubei",
                                  param_list$p_report_source[1],
                                  param_list$p_report_source[2]),
           # For first pass, reporting rate is just Hubei/not Hubei
           days_per_t=param_list$delta # ~ delta: days per time period
    ) %>%
    dplyr::filter(t<=as.Date(last_date)) %>%
    dplyr::mutate(source = as.character(source),
           destination = as.character(destination))
  
  ## Filter to sources with cases -- to speed it up
  source_w_cases <- input_data %>%
    dplyr::filter(!duplicated(paste0(source, t))) %>%
    dplyr::group_by(source) %>%
    dplyr::summarise(cum_cases = sum(cases_incid, na.rm=TRUE)) %>%
    dplyr::filter(cum_cases>0)
  input_data <- input_data %>%
    dplyr::filter(source %in% source_w_cases$source)
  travel_data_monthly <- travel_data_monthly %>%
    dplyr::filter(source %in% source_w_cases$source)
  travel_data_daily <- travel_data_daily %>%
    dplyr::filter(source %in% source_w_cases$source)
  
  
  # save the data that we will pass to the model
  readr::write_csv(input_data, file.path(output_dir, "input_data.csv"))
  readr::write_csv(travel_data_monthly, file.path(output_dir, "travel_data_monthly.csv"))
  readr::write_csv(travel_mean, file.path(output_dir, "travel_mean.csv"))
  readr::write_csv(travel_data_daily, file.path(output_dir, "travel_data_daily.csv"))
  
  print(paste0("Input and Travel data setup successfully and saved in ", output_dir, "."))

  
  


```


## Run the Simulations

```{r, warning=FALSE, message=FALSE}

sim_res <- covidImportation::run_importations(
                             n_sim=1000,
                             cores=5,
                             get_detection_time=FALSE,
                             travel_dispersion=3,
                             allow_travel_variance=FALSE,
                             print_progress=TRUE,
                             output_dir = file.path("importation",regioncode),
                             param_list=list(incub_mean_log=log(5.89),
                                             incub_sd_log=log(1.74),
                                             inf_period_nohosp_mean=15,
                                             inf_period_nohosp_sd=5,
                                             inf_period_hosp_shape=0.75,
                                             inf_period_hosp_scale=5.367,
                                             p_report_source=c(0.05, 0.25)))

```
    
    
## Get shapefiles
# DONT NEED THIS HERE

```{r, warning=FALSE, message=FALSE, eval=FALSE}

tidycensus::census_api_key(key="[ENTER KEY HERE]")

countypops <- get_county_pops(states_of_interest=states_of_interest, 
                            regioncode = regioncode, 
                            yr=2018, 
                            local_dir="data/", 
                            write_county_shapefiles=TRUE)


```



## Distribute the Simulated Importations into Airports to Counties

```{r, warning=FALSE, message=FALSE, eval=FALSE}

tidycensus::census_api_key(key="[ENTER KEY HERE]")

run_full_distrib_imports(states_of_interest=states_of_interest,
                                     regioncode=regioncode,
                                     yr=2018,
                                     mean_travel_file = file.path("importation", regioncode, "travel_mean.csv"),
                                     travelers_threshold=10000,
                                     airport_cluster_threshold=80,
                                     shapefile_path = NULL,
                                     model_output_dir = file.path("importation", regioncode),
                                     local_dir="data/",
                                     plot=FALSE,
                                     cores=5,
                                     n_sim=100)

```


```{r}


FIPS_data <- readr::read_csv("data_other/us_terr_airports.csv") #%>% filter(state!="PR")

library(doParallel)
cores=5
print(paste0("Making a cluster of ", cores," for parallelization."))
cl <- parallel::makeCluster(cores)
doParallel::registerDoParallel(cl)
n_sim=1000

yr=2018
airport_cluster_threshold=80
mean_travel_file = file.path("data", "travel_mean.csv")
shapefile_path = NULL ## This is no longer needed but left to not break it.
model_output_dir = file.path("importation")
local_dir="data/"


# Run the foreach loop to estimate importations for n simulations
foreach(n=seq_len(n_sim),
        .packages=c("dplyr","tidyr","readr")) %dopar% {
    
    imports_sim <- readr::read_csv(file.path(model_output_dir, regioncode, paste0("imports_sim",n,".csv"))) %>%
        rename(amount = this.sim, date=t)
    imports_sim <- left_join(imports_sim, FIPS_data, by="destination") %>%
        filter(state !="PR") %>% 
        group_by(date, FIPS) %>%
        summarise(amount = sum(amount)) %>%
        rename(place = FIPS) %>%
        select(place, date, amount)
    
    write_csv(imports_sim, file.path(model_output_dir, regioncode, paste0("importation_",n,".csv")))
        }

parallel::stopCluster(cl)




```








