---
title: "nCoV Importation - Shenzhen"
author: "Shaun Truelove"
email: "shauntruelove@jhu.edu"
date: "`r Sys.time()`"
geometry: margin=.5in
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}

options(scipen = 999)
knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = TRUE)

if(!require('knitr')) install.packages('knitr'); library(knitr)
if(!require('tidyverse')) install.packages('tidyverse'); library(tidyverse)
if(!require('gridExtra')) install.packages('gridExtra'); library(gridExtra)
if(!require('grid')) install.packages('grid'); library(grid)
if(!require('viridis')) install.packages('viridis'); library(viridis)
select <- dplyr::select


source("R/DataLoadUtils.r")
source("R/BasicEpiAnalyses.r")


```

This markdown document details the full analysis process for estimating the 2019-nCoV importation risk and numbers into the U.S. from other countries. This work is an adaptation of work done for nCoV importation for the manuscript titled "Epidemics, Air Travel, and Elimination in a Globalized World: The Case of Measles". The R scripts required are noted/sourced below, though most are not run with this rmd. To do the full analysis, data from IATA on full trip iteneraries is required, but this is not publicly available, so we only include the summarized/aggregated results of the model.
    
# MODEL SETUP

### Project Options

```{r projoptions, echo=TRUE}
# Analysis Options
project_name <- "shenzhen_import"
version <- "All_timevar"
batch <- "1st"
n_sim <- 2000
```

    
    
## nCoV Characteristics

### Incubation Period
These estimates are from Lauer et al. 2020 (https://www.medrxiv.org/content/10.1101/2020.02.02.20020016v1).
```{r incub, echo=TRUE}

# Incubation
# mean: 5.2 days (95% CI 2.5-10.5) 
incub_mean_log <- log(5.2)
incub_sd_log   <- log(1.43)

samp_incub <- exp(rnorm(10000, incub_mean_log, incub_sd_log))
quantile(samp_incub, probs=c(.5))
quantile(samp_incub, probs=c(0.025, 0.975))

```

### Infectious Period
We are using different infectious periods for those detected and hospitalized and those not.
These estimates are based off ...

#### Infectious period - not hospitalized
```{r infectperiod_nohosp, echo=TRUE}
inf_period_nohosp_mean <- 7  # needs revision
inf_period_nohosp_sd   <- 4

#inf_nohosp <- exp(MCMCglmm::rtnorm(10000, log(inf_period_nohosp_mean), log(inf_period_nohosp_sd), lower=0))
inf_nohosp <- (MCMCglmm::rtnorm(10000, inf_period_nohosp_mean, inf_period_nohosp_sd, lower=0))

quantile(inf_nohosp, probs=c(0.025, .5, 0.975))
print(paste0("Mean time to recovery: ", round(mean(inf_nohosp),1), " days"))

```

#### Infectious period - hospitalized (time to hospitalization)
```{r infectperiod_hosp, echo=TRUE}

inf_period_hosp_mean_log <- 1.23  
inf_period_hosp_sd_log <- 0.79

inf_hosp <- rlnorm(1000, meanlog=inf_period_hosp_mean_log, sdlog=inf_period_hosp_sd_log)

quantile(inf_hosp, probs=c(0.025, .5,  0.975))
print(paste0("Mean time to hospitalization: ", round(mean(inf_hosp),1), " days"))

```


```{r plot_params, echo=FALSE}

par(mfrow=c(2,2))

plot(density(samp_incub), type="l", main="Incubation Period", xlab="Days")

plot(density(inf_nohosp), type="l", col="blue", ylim=c(0,.3), xlim=c(0, 30), 
     main="Time to recovery", xlab="days")
plot(density(inf_hosp), type="l", col="red", ylim=c(0,.35), xlim=c(0, 30), 
        main="Time to hospitalization", xlab="days")

par(mfrow=c(1,1))

```



### Proportion Cases Reported
Currently the model is set up to take both time- and location-varying proportion of cases reported. For now, we are using 5% reported for Hubei, and 25% reported elsewhere, constant over time.
```{r}
p_report_source <- c(0.05, 0.25)

```

  

# DATA

Here we describe where we get data and do necessary formatting and cleaning of the data.

Three major sets of data are required for this importation estimation model: incidence in source locations, mobility or travel volume from source locations in to destination locations, and population of source locations. Data processing is done within separate R scripts for each. These each are setup to process and save the files in the correct formats for running the model.
**To run these source scripts within this Rmd file, change `eval=FALSE` to `eval=TRUE`. 


## Incidence Data
Process and plot the incidence data. We will use a spline fitting to estimate te incidence from reported case counts, as case reporting is unreliable.
    
```{r incid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, fig.height=4}

## If want to update the linelist thats in a google doc, 
##  change this to TRUE (in china, keep as FALSE)
get_googledocs_data <- FALSE

# Pull and save data from github
pull_JHUCSSE_github_data()

## If you need to run full data processing, run this:
source("R/ncov_incidence_data_processing.R")



## Get case count data (from JHU CSSE's github), and estimate incidence using spline fits.
incid_data <- get_incid_data(pull_data = FALSE, 
                             first_date = ISOdate(2019,12,1), last_date=ISOdate(2020,2,16), 
                             tol=10)

## Plot current confirmed cases in China
conf_cases <- read_JHUCSSE_cases(last_time = Sys.time(), append_wiki=TRUE) %>% 
  filter(Country_Region == "Mainland China" | Province_State %in% c("Hong Kong", "Macau", "Taiwan")) %>%
  mutate(t = as.Date(Update)) %>% arrange(Province_State, Country_Region, Update) %>%
  group_by(Country_Region, Province_State) %>% mutate(Incidence = diff(c(0, Confirmed), na.rm=TRUE)) %>% ungroup() %>%
  group_by(Country_Region, Province_State, t) %>% summarise(Incidence = sum(Incidence, na.rm = TRUE))

conf_cases <- conf_cases %>% filter(t >= as.Date("2020-01-15"))
t_values <- as.character(sort(conf_cases$t))

p <- ggplot(data=conf_cases, aes(x=as.Date(t), y=Incidence, fill=Province_State)) + 
        geom_bar(position="stack", stat="identity", color="black") +
        ylab("Confirmed nCoV cases (n)") +
        theme(axis.line = element_line(colour = "black"),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
              panel.border = element_blank(), panel.background=element_blank(),
              axis.text.y = element_text(size=9), 
              axis.text.x = element_text(size=9, angle=45, hjust = 1),
              axis.title.x = element_blank(), legend.title=element_blank(), 
              legend.text=element_text(size=7), legend.key.size=unit(6, "pt"),
              #legend.margin=margin(0,0,0,0), legend.box.margin=margin(20,0,-20,-20),
              legend.background = element_blank(), legend.box.background = element_blank(),
              plot.title = element_text(size=8, face="bold", hjust = 0.025),
              plot.margin = unit(c(0.5,.25,0.25,0.25), "cm")) +
        guides(fill=guide_legend(ncol=1))
plot(p)
```




## Reported Importation data    

```{r reported_cases, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4}
#These get cleaned in `source("R/ncov_incidence_data_processing.R")`.

# Reported from CDC
shen_cases <- read_csv("data/shenzhen_data/shenzhen_case_counts.csv")
shen_cases <- shen_cases %>% mutate(cum_cases = cumsum(count))
# From Linelists
ll_data <- read_csv("data/linelist_current.csv")
shen_rows <- apply(ll_data, 1, FUN=function(x) sum(grepl("shenzhen", x, ignore.case = TRUE)))>0
ll_data_shenzhen <- ll_data[shen_rows, ]
shen_data_aggr <- ll_data_shenzhen %>% count(date_confirmation)
rm(ll_data, ll_data_shenzhen, shen_rows)

shen_counts <- full_join(shen_cases %>% rename(count_CDC = count),
                         shen_data_aggr %>% rename(count_ll = n, date=date_confirmation),
                         by=c("date"="date"))
shen_counts[is.na(shen_counts)] <- 0


# Plot the epi curve of these
p_incid_shen <- ggplot(shen_counts %>% mutate(label1 = "CDC counts", label2 = "Linelist counts"), 
                     aes(x=date, y=count_CDC, color=label1, group=label1)) +
                  geom_bar(stat="identity", fill="darkblue") +
                  geom_point(aes(x=date,y=count_ll, shape=label2), color="maroon", size=2) +
                  scale_color_manual(values=c(NA, "maroon")) +
                  ylab("Reported Cases") + ggtitle("Shenzhen Reported nCoV Cases") + 
                  xlab(NULL) +
                  theme_classic() +
                  theme(legend.position = c(0.05, 1), legend.justification = c(0, 1),
                        legend.background = element_blank(),
                        legend.title = element_blank())
                

# Plot cumulative cases in Shenzhen
# Get cumulative case counts in each
shen_counts <- shen_counts %>% arrange(date) %>% mutate(cum_CDC = cumsum(count_CDC), cum_ll = cumsum(count_ll))

p_cum_shen <- ggplot(shen_counts %>% mutate(label1 = "CDC cumulative", label2 = "Linelist cumulative"), 
       aes(x=date, y=cum_CDC, color=label1, group=label1)) +
    geom_bar(stat="identity", fill="darkblue") +
    geom_point(aes(x=date,y=cum_ll, shape=label2), color="maroon", size=2) +
    scale_color_manual(values=c(NA, "maroon")) +
    ylab("Cumulative Reported Cases") + ggtitle("Shenzhen Cumulative Reported nCoV Cases") + xlab(NULL) +
    theme_classic() +
    theme(legend.position = c(0.05, 1), legend.justification = c(0, 1),
          legend.background = element_blank(),
          legend.title = element_blank())

p1 <- ggplot_gtable(ggplot_build(p_incid_shen))
p2 <- ggplot_gtable(ggplot_build(p_cum_shen))
maxWidth = grid::unit.pmax(p1$widths[2:5], p2$widths[2:5])
p1$widths[2:5] <- as.list(maxWidth)
p2$widths[2:5] <- as.list(maxWidth)
gridExtra::grid.arrange(p1, p2, nrow=2)


```

  

## Mobility Data    

```{r travel, echo=FALSE, message=FALSE, eval=TRUE, fig.height=5}

source("R/setup_travel_data.R")
travel_data <- read_csv("data/shenzhen_travel_data.csv")

# Plot Travelers

t_ <- unique(travel_data$t)
source_levels <- c("Guangdong", unique((travel_data %>% filter(source!="Guangdong"))$source))

all_travel <- ggplot(travel_data %>% mutate(source = factor(source, levels=source_levels)), 
                  aes(t, travelers, group=source, fill=source)) + 
                    guides(fill=guide_legend(title = element_blank(), ncol=1)) +
                    geom_area(color='black') +
                    theme_classic() +
                    theme(axis.line = element_line(colour = "black"),
                      axis.text.x = element_text(angle=45, hjust=1),
                      axis.title.x = element_blank(), legend.title=element_blank(), 
                      legend.margin=margin(0,0,0,0), legend.box.margin=margin(0,0,0,0),
                      legend.text=element_text(size=7), legend.key.size=unit(6, "pt"), 
                      legend.background = element_blank(), legend.box.background = element_blank(),
                      plot.title = element_text(size=8, face="bold", hjust = 0.025),
                      plot.margin = unit(c(0.25,.25,0.25,0.25), "cm"))

# Hubei only
hubei_travel <- ggplot(travel_data %>% filter(source=="Hubei"), 
                       aes(t, travelers, group=source, fill=source)) + 
                      guides(fill=guide_legend(title = element_blank(), ncol=1)) +
                      theme_classic() +
                      theme(axis.text.x = element_text(angle = 45, hjust = 1),
                            legend.text=element_text(size=7), legend.key.size=unit(6, "pt"),
                            legend.margin=margin(0,15,0,5), legend.box.margin=margin(0,0,0,0),
                            plot.margin = unit(c(0.25,.25,0.25,0.25), "cm")) +
                  geom_area(color='black') + viridis::scale_fill_viridis(discrete = T) +
                  xlab(NULL)


p1 <- ggplot_gtable(ggplot_build(all_travel))
p2 <- ggplot_gtable(ggplot_build(hubei_travel))
maxWidth = grid::unit.pmax(p1$widths[2:5], p2$widths[2:5])
p1$widths[2:5] <- as.list(maxWidth)
p2$widths[2:5] <- as.list(maxWidth)
gridExtra::grid.arrange(p1, p2, nrow=2, heights=c(2.5,1))

```

We see that travel from Hubei had a massive spike on 27 Jan, 4 days after the lockdown on 23 Jan. Whether this was from people still in transit, or people fleeing before restrictions were actually in place is unclear. Overall, travel from Hubei is relatively minimal compared to that from other locations.
    
    
## Population Data

```{r pop, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
source("R/setup_pop_data.R")
```



  
# MODEL

```{r model, echo=TRUE, eval=F, message=FALSE, warning=FALSE}
version="All_timevar"
batch <- "1st"
# shift incidence back by 4 days (midpoint of incubation period (=2) + time to detect/seek care (=2))
shift_incid_days <- -5 
source("R/covid_importation_model_time&loc.R")

```
    
    
# RESULTS

### Analysis of Model Simulations     
Clean, aggregate, and produce summary stats for importation simulation results.

```{r model_res, echo=FALSE, eval=F, message=FALSE, warning=FALSE}

# Time and location varying
version="All_timevar"
batch <- "1st"

t.start <- proc.time() # start timer to measure this
source("R/importation_analysis.R")
source("R/importation_analysis_detection.R")

# print time required
print(paste0('Analysis required ', round(as.list(proc.time() - t.start)$elapsed/60, 3), ' minutes'))

```



## Figure 1. Heatmaps of Risk of Importation From other Provinces, All Provinces.
    
```{r plot_heatmapsA, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}

# Requires: `R/ggheat_func_source.R` and `R/heatmaps_exportation.R`    

version="All_timevar"
batch <- "1st"

source("R/heatmaps_exportation.R")

gridExtra::grid.arrange(p_rr, p_prob, p_mean, 
                        layout_matrix=rbind(c(1,1,1,1), c(1,1,1,1), c(1,1,1,1), c(1,1,1,1),
                                            c(2,2,3,3), c(2,2,3,3), c(2,2,3,3)))
```

We can see very clearly that the majority of risk is coming from Hubei by far, though importations may be coming from other provinces as the outbreak grows. 



## Figure 2. Estimated Importations into Shenzhen, All Provinces.

```{r plot_importsA, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6}

version="All_timevar"
batch <- "1st"
source("R/plot_importation_estimates.R")

# Plot them
gridExtra::grid.arrange(p_total, p_cum, nrow=2)

```


    
```{r plot_importsA2, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4.25}

# Plot them
plot(p_stacked_barchart)

```
    
    
    
## Figure 3. Detected Estimated Importations into Shenzhen, All Provinces.

```{r plot_importsB, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6}

version="All_timevar"
batch <- "1st"
t_limits <- as.Date(c("2020-01-18","2020-02-17"))

source("R/plot_importation_estimates_detection.R")

# Plot these
gridExtra::grid.arrange(p_incid_shen_inclest, p_cum_shen_inclest, nrow=2)

```

The estimated importations line up well with detection, but about 2-3 days before, which is likely explained by delays to confirmation and reporting.
Detection time is based on time from onset to hospitalization.

    
```{r plot_importsB2, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}

plot(p_stacked_barchart)

```

These Estimates of detected importations are currently only valid until around Jan 30, because of the need for future case projections. The model currently uses confirmed cases to make estimates of previous infections, assuming current confirmations represent cases that were infected in the past. The time from infection to confirmation is a function of the incubation period, time to seek care, and time to be confirmed. The next update will use near-term case projections to estimate the number of infected individuals potentially traveling in current and near future time.
