---
title: "R Notebook"
output: html_notebook
---


Use Travel Data from past 3 years to estimate travel during 2020
We will smooth across the years to account for random fluctuations


```{r setup}
options(scipen = 999)
#knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = TRUE)

if(!require('knitr')) install.packages('knitr'); library(knitr)
if(!require('tidyverse')) install.packages('tidyverse'); library(tidyverse)
if(!require('gridExtra')) install.packages('gridExtra'); library(gridExtra)
if(!require('grid')) install.packages('grid'); library(grid)
if(!require('viridis')) install.packages('viridis'); library(viridis)

```


## DATA

```{r data, message=FALSE}
# DATA --------------------------------------------------------------------

## Full travel data
#travel_data <- read(file.path("data", "complete_OAG_data.csv"))

## California travel data (we will start with this)
travel_data <- read_csv("data/cali_oag_20172019.csv")
travel_data <- travel_data %>% rename(dest_city=`Arr City Name`, dest_airport=`Arr Airport Code`,dest_state=`Arr State Code`, dest_country=`Arr Country Code`,
                                      source_city=`Dep City Name`, source_airport=`Dep Airport Code`, source_state=`Dep State Code`, source_country=`Dep Country Code`,
                                      travelers = `Total Est. Pax`, yr_month=`Time Series`) %>%
    mutate(source=source_airport, destination=dest_airport) %>%
    mutate(t_year = substr(yr_month, 1,4),
           t_month = substr(yr_month, 5,6),
           yr_month = factor(yr_month, levels=sort(unique(yr_month)), ordered = TRUE))

```


### SFO Testing
Lets work on a single destination (SFO)

```{r}
sfo_data <- travel_data %>% filter(dest_airport == "SFO")
```

Plot a couple source locations to look at the variations
```{r}

#sample of sources
samp <- sample(unique(sfo_data$source_airport), 20)
ggplot(sfo_data %>% filter(source_airport %in% samp), aes(t_month, travelers, group=t_year, color=t_year)) +
    geom_line() + geom_point() +
    facet_wrap(vars(source_airport), nrow=5, ncol=4)

```

Let's make sure there are no weird things for low, mid, and high travel sources.

```{r}
## Get annual total and mean annual total
sfo_data <- sfo_data %>% group_by(source_airport, t_year) %>% mutate(year_total=sum(travelers)) %>% 
    group_by(source_airport) %>% mutate(yeartot_mean = round(mean(year_total))) %>% ungroup()

# total number of sources
length(unique(sfo_data$source_airport))
hist((sfo_data %>% filter(!duplicated(source_airport)))$yeartot_mean, breaks=100, main="Distribution of total annual travelers")
hist((sfo_data %>% filter(!duplicated(source_airport)))$yeartot_mean,  breaks=10000, xlim=c(0,10000),main="Distribution of total annual travelers")
hist((sfo_data %>% filter(!duplicated(source_airport)))$yeartot_mean,  breaks=10000, xlim=c(0,1000),main="Distribution of total annual travelers")
quantile((sfo_data %>% filter(!duplicated(source_airport)))$yeartot_mean, probs = seq(0,1,.1))


#sample of sources 
sampvhigh <- sample(unique((sfo_data %>% filter(yeartot_mean>100000))$source_airport), 20)
samphigh <- sample(unique((sfo_data %>% filter(yeartot_mean<=100000 & yeartot_mean>10000))$source_airport), 20)
sampmid <- sample(unique((sfo_data %>% filter(yeartot_mean<=10000 & yeartot_mean>1000))$source_airport), 20)
samplow <- sample(unique((sfo_data %>% filter(yeartot_mean<=1000))$source_airport), 20)


# Very high travel
ggplot(sfo_data %>% filter(source_airport %in% sampvhigh), aes(t_month, travelers, group=t_year, color=t_year)) +
    geom_line() + geom_point() +
    facet_wrap(vars(source_airport), nrow=5, ncol=4)

# High travel
ggplot(sfo_data %>% filter(source_airport %in% samphigh), aes(t_month, travelers, group=t_year, color=t_year)) +
    geom_line() + geom_point() +
    facet_wrap(vars(source_airport), nrow=5, ncol=4)

# Mid travel
ggplot(sfo_data %>% filter(source_airport %in% sampmid), aes(t_month, travelers, group=t_year, color=t_year)) +
    geom_line() + geom_point() +
    facet_wrap(vars(source_airport), nrow=5, ncol=4)

# Low travel
ggplot(sfo_data %>% filter(source_airport %in% samplow), aes(t_month, travelers, group=t_year, color=t_year)) +
    geom_line() + geom_point() +
    facet_wrap(vars(source_airport), nrow=5, ncol=4)



```

Generally speaking, flight volume looks relatively consistent across the various levels. We will do estimation using the 3 years.




```{r}


m1 <- glm(travelers ~ t_month*source_airport, family="poisson", data=sfo_data) #data=sfo_data %>% filter(source_airport=="WUH"))




```






